---
title: "EDA"
output:
  html_document:
    toc: true
    toc_depth: 2
---

<style type="text/css">
  body{
  font-size: 16pt;
}
</style>

```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE,
                      message = FALSE)

```


```{r load_libraries}

devtools::load_all()

library(tidyverse)

library(tidymodels)

library(embed)

library(corrr)

library(ggcorrplot)

library(glue)

```

```{r set_params}

professions = c(
  "manager",
  "business",
  "financialop",
  "computer",
  "architect",
  "scientist",
  "socialworker",
  "postseceduc",
  "legaleduc",
  "artist",
  "lawyerphysician",
  "healthcare",
  "healthsupport",
  "protective",
  "foodcare",
  "building",
  "sales",
  "officeadmin",
  "farmer",
  "constructextractinstall",
  "production",
  "transport"
)



```

```{r load_data}

data(train_set)

train_set_transformed = train_set %>%
  rename(ID = 1) %>%
  pivot_longer(cols = all_of(professions), names_to = "profession") %>%
  filter(value == 1) %>%
  select(-value) %>%
  mutate(profession = factor(profession))%>% 
  rowwise() %>% 
  mutate(white = 1 - sum(c_across(c("black","hisp","otherrace")))) %>% 
  ungroup() %>% 
  pivot_longer(cols = c(c("black","hisp","otherrace","white")),
               names_to = "race") %>% 
  filter(value == 1) %>% 
  select(-value) %>% 
  rowwise() %>% 
  mutate(west = 1 - sum(c_across(c("northeast","northcentral","south")))) %>% 
  ungroup() %>% 
  pivot_longer(cols = c(c("northeast","northcentral","south","west")),
               names_to = "area") %>%
  filter(value == 1) %>%
  select(-value) %>%
  mutate(male = 1 - female) %>% 
   pivot_longer(cols = c(c("male","female")),
               names_to = "sex") %>%
  filter(value == 1) %>%
  select(-value) %>% 
  mutate(across(where(is.character),as.factor)) %>% 
  mutate(across(where(~length(unique(.)) <= 2), as.factor)) %>% 
  identity()


# data(test_set)
# 
# test_set = test_set %>% 
#   rename("ID" = 1)

```

```{r recipe_preprocessing}

preprocess_recipe = recipe(lnwage ~ ., train_set_transformed) %>% 
  update_role(ID,new_role = "id variable") %>% 
  step_rm(ends_with("sq"))%>%
  step_YeoJohnson(all_numeric(),-ID,-lnwage) %>% 
  step_mutate(part_time_share = expp / (expp + expf + 0.01)) %>% 
  step_mutate(exp_sq = (expp + expf) ^ 2) %>% 
  step_lencode_glm(c(profession,race,sex, area),outcome = vars(lnwage)) %>% 
  step_interact(~edyrs:race) %>%
  step_interact(~sex:race)


train_set_baked = preprocess_recipe %>% 
  prep() %>% 
  bake(train_set_transformed)
  

```


## Outliers

```{r outliers}

train_set_transformed %>% 
  select(where(~ is.numeric(.x) && length(unique(.x)) > 2),-ID) %>% 
  pivot_longer(everything()) %>% 
  ggplot(aes(y = value)) + 
  geom_boxplot() + 
  facet_wrap(~name, scales = "free")

```

We can see that there are outliers in all the numeric features.
For the squared features I'll try the Yeo-Johnson transformation to deal
with skewness.


```{r plot_outliers_after_transformation}

train_set_baked %>% 
  select(where(~ is.numeric(.x) && length(unique(.x)) > 2),-ID) %>% 
  pivot_longer(everything()) %>% 
  ggplot(aes(y = value)) + 
  geom_boxplot() + 
  facet_wrap(~name, scales = "free")

```

## Numeric features


```{r dist_of_numeric_features}

train_set_baked %>% 
  select(where(~ is.numeric(.x) && length(unique(.x)) > 2)) %>% 
  pivot_longer(cols = -ID) %>% 
  ggplot(aes(x = value)) + 
  geom_histogram() + 
  facet_wrap(~name, scales = "free") + 
  xlab(NULL) + ylab(NULL)
  

```

When looking at the correlations of the numeric features with lnwage we see that education and full time experience are positively correlated with lnwage. Part time is negatively correlated.

```{r correlations_of_numeric_features}

train_set_baked %>% 
  select(-ID) %>% 
  select(where(~ is.numeric(.x) && length(unique(.x)) > 2)) %>% 
  correlate() %>% 
  stretch() %>% 
  rename(feature = y) %>% 
  filter(complete.cases(.)) %>% 
  filter(x == "lnwage") %>% 
  ggplot(aes(x = reorder(feature,r), y = r)) + 
  geom_col() + 
  coord_flip() + 
  xlab(NULL) + ylab(NULL) + 
  ggtitle("Correlations of numeric features with lnwage")

```


```{r scatterplot_of_numeric_features}

train_set_baked %>% 
  select(where(~ is.numeric(.x) && length(unique(.x)) > 2), lnwage,-ID) %>% 
  pivot_longer(-lnwage) %>% 
  ggplot(aes(x = value, y = lnwage)) + 
  geom_point() + 
  geom_smooth(method = "glm") + 
  facet_wrap(~name, scales = "free") + 
  xlab(NULL) + ylab(NULL) + 
  ggtitle("Scaterrplot of lnwage and numeric features")

```


## Categorical features

There are many features that are the results of one hot encoding of the categorical features. Perhaps we can try another encoding and so I collect back the categories in area, race and profession. Both area and race are missing the reference category (west-northwest for area and white for race)


```{r boxplot_by_target, fig.height=7}

train_set_transformed %>% 
  select(where(is.factor), lnwage) %>% 
  pivot_longer(-lnwage) %>% 
  ggplot(aes(x = value, y = lnwage)) + 
  geom_boxplot() + 
  facet_wrap(~name, scales = "free") + 
  xlab(NULL) + ylab(NULL) + ggtitle("lnwage distribution by cat features")



```

### Area

South is the most low paying and the most frequent area


```{r target_distribution_by_area}

train_set_baked %>% 
  select(lnwage, area) %>% 
  ggplot(aes(x = reorder(area,lnwage,FUN = mean), y = lnwage)) + 
  geom_boxplot() + 
  coord_flip() + 
  xlab(NULL) + ylab(NULL) + ggtitle("Average lnwage by area")

```

```{r area_distribution}

train_set_baked %>% 
  count(area) %>% 
  mutate(n = n/sum(n)) %>% 
  ggplot(aes(x = reorder(area,n), y = n)) + 
  geom_col() + 
  coord_flip() + 
  scale_y_continuous(labels = scales::percent_format()) + 
  xlab(NULL) + ylab(NULL) + ggtitle("Area distribution")
  

```



### Profession

In general the data looks good (computer, business and finance making more than social worker and building) but there are "outliers" such as artist (at the top) or scientist (at the middle).Checking out the distribution of lnwage by profession we can see that there is one rich artist that stands out. One way of dealing with it is to try to encode with median lnwage by profession


```{r target_distribution_by_profession}

train_set_transformed %>% 
  select(lnwage, profession) %>% 
  ggplot(aes(x = reorder(profession,lnwage,FUN = mean), y = lnwage)) + 
  geom_boxplot() + 
  coord_flip() + 
  xlab(NULL) + ylab(NULL) + ggtitle("Average lnwage by profession")

```

```{r profession_distribution}

train_set_transformed %>% 
  count(profession) %>% 
  mutate(n = n/sum(n)) %>% 
  ggplot(aes(x = reorder(profession,n), y = n, fill = (n <= 0.025))) + 
  geom_col() + 
  coord_flip() + 
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(fill = "Rare profession (less than 2.5%)") + 
  xlab(NULL) + ylab(NULL) + ggtitle("Profession distribution")
  

```


We can see that some professions are relatively rare so we may try to drop
them all in one category (other)


```{r target_distribution_by_profession_after_lumping}

train_set_baked %>% 
  select(lnwage, profession) %>% 
  ggplot(aes(x = reorder(profession,lnwage,FUN = mean), y = lnwage,
             fill = (profession == "other"))) + 
  geom_boxplot() + 
  coord_flip() + 
  scale_fill_manual(values = c("white","lightblue")) + 
  xlab(NULL) + ylab(NULL) + ggtitle("Average lnwage by profession") + 
  theme(legend.position = "none")

```



### Race


```{r target_distribution_by_race}

train_set_transformed %>% 
  select(lnwage, race) %>% 
  ggplot(aes(x = reorder(race,lnwage,FUN = mean), y = lnwage)) + 
  geom_boxplot() + 
  coord_flip() + 
  xlab(NULL) + ylab(NULL) + ggtitle("Average lnwage by race")

```

```{r race_distribution}

train_set_transformed %>% 
  count(race) %>% 
  mutate(n = n/sum(n)) %>% 
  ggplot(aes(x = reorder(race,n), y = n,fill = (n <= 0.05))) + 
  geom_col() + 
  coord_flip() + 
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(fill = "Rare race (less than 5%)") + 
  xlab(NULL) + ylab(NULL) + ggtitle("Race distribution")
  

```



##Interactions

```{r check_for_race_edyrs_interaction}

train_set_baked %>%
  ggplot(aes(x = edyrs, y = lnwage, color = race)) + 
  geom_smooth(method = "lm", se = FALSE)

```

There seems to be an interaction between race and education - the effect of education on wage changes with race.

```{r add_interaction_between_edyrs_and_race}

preprocess_recipe = preprocess_recipe 

train_set_baked = preprocess_recipe %>% 
  prep() %>% 
  bake(train_set_transformed)

```

```{r check_for_area_edyrs_interaction}

train_set_transformed %>% 
  ggplot(aes(x = edyrs, y = lnwage, color = area)) + 
  geom_smooth(method = "lm", se = FALSE)

```

Looking at an interaction between area and education gives nothing.



# Summary
* Reduce skewness by winsorizing
* Reverse OHE in area, race and profession
* For profession : 
  * try (lnwage) embeddings
  * lump rare categories together  





